---
title: "Linear"
output: html_document
date: "2024-01-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data

The given data is sampled from a bivariate normal distribution. 

```{r}
x<-c(66,69,65,63,65,64,62,65,64,59,62)
y<-c(72,71,66,67,70,68,71,70,73,65,66)
```

## Linear Regression

We then can compute the sample coefficient of determination:

```{r pressure, echo=FALSE}
r=cor(x,y)
rsquare=r^2
```
## Prediction

We can use this data in order to determine if there is enough evidence that knowing $x$ will help one predict $y$.

```{r}
model=lm(y~x)
summary(model)
```
## Summary

Then, since $p=0.07442>\alpha=0.05$ we cannot say for certain that knowing $x$ helps one predict $y.$

Using this result, we can construct a $90 \%$ confidence interval for the slope of the population regression line for predicting $y$ from $x$.

In order to construct a confidence interval, we will need the formula: 
 \[\hat{\beta_1} \pm qt(0.95, \ df=9)*se(\hat{\beta_1})\]
 where $\hat{\beta_1}$ is the estimated slope of the regression line and $se$ is the standard error. For $s_x=sd(x)=2.57$, $s_y=sd(y)=2.72$, $n=11$, and $r=0.558$, the standard error is,
 \begin{align*}
     se(\hat{\beta_1}) &= \frac{s_y}{s_x}\sqrt{\frac{1-r^2}{n-2}} \\
     &=\frac{2.72}{2.57}\sqrt{\frac{1-0.311}{11-2}} \\
     &= 0.2928
 \end{align*}
 
 Then, 
 $$\hat{\beta_1} = r \frac{s_y}{s_x} = 0.5909$$
 Giving us,
 \[\hat{\beta_1} \pm qt(0.95, \ df=9)*se(\hat{\beta_1}) = 0.5909 \pm 0.5367354864 \]
 Therefore, a 0.90-level confidence interval for the slope of the population regression line for predicting $y$ from $x$ is (0.0542,1.128). 
